{Supporting Information for 'Efficient Detection and Classification of Epigenomic Changes Under Multiple Conditions' by Pedro L. Baldoni$^*$, Naim U. Rashid$^*$, and Joseph G. Ibrahim$^*$}

### 2021-02-18 Latex/appendix/webappendix.tex
Adding latex files from appendix & paper

\documentclass[12pt,oneside]{book}

@@

\if0\blind{
\date{%
\today
}
\title{{Supporting Information for 'Efficient Detection and Classification of Epigenomic Changes Under Multiple Conditions'}}
} \fi

\begin{document}
\maketitle

@@

For csaw, we used the following setup:
<<echo=TRUE,eval=FALSE,cache=TRUE>>=

# List of bam files
bam.files = list.files(path=paste0(tmpdir,'/chip'),pattern='*.bam$',
                       full.names=T)
bam.files

# Design matrix
design <- model.matrix(~factor(cell.type))
colnames(design) <- c("intercept", "cell.type")
design

# Parameters (PCR duplicates already removed and quality score filtered)
param <- readParam(dedup = F)
param

# Estimating the average fragment length (rescaling all to 200bp)
x = lapply(bam.files,correlateReads,param=param,max.dist=250)
multi.frag.lens = list(unlist(lapply(x,maximizeCcf)),200)
multi.frag.lens

# Counting reads (for a window size of 250bp, for instance)
data <- windowCounts(bam.files,width = 250,ext = multi.frag.lens,
                     param = param,filter = 20)
data

# Filtering data
data.large <- windowCounts(bam.files,width=2500,bin=T,param=param)

bin.ab <- scaledAverage(data.large, scale=median(getWidths(data.large))/
                          median(getWidths(data)))

threshold <- median(bin.ab) + log2(2)

keep.global <- aveLogCPM(asDGEList(data)) >  threshold

sum(keep.global)

# Creating filtered data
filtered.data <- data[keep.global,]

# Testing for DB (assuming composition bias is negligble,
# i.e. cell lines should exhibit a balanced number of DB regions)

y <- DGEList(assay(filtered.data), lib.size = filtered.data$totals)
y$samples$norm.factors <- 1 
y$offset <- NULL
y <- estimateDisp(y, design)
fit <- glmQLFit(y, design, robust = TRUE)
out <- glmQLFTest(fit,contrast = contrast)
tabres <- topTags(out, nrow(out))$table
tabres <- tabres[order(as.integer(rownames(tabres))),]

merged <- mergeWindows(rowRanges(filtered.data), tol=tol,
                       max.width=max.width)
tabneg <- combineTests(merged$id, tabres)
@

>>>>

\documentclass[12pt,oneside]{book}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

\if0\blind{
\date{%
\today
}
\title{{Supporting Information for 'Efficient Detection and Classification of Epigenomic Changes Under Multiple Conditions'}}
} \fi
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle

@@

For csaw, we used the following setup:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# List of bam files}
\hlstd{bam.files} \hlkwb{=} \hlkwd{list.files}\hlstd{(}\hlkwc{path}\hlstd{=}\hlkwd{paste0}\hlstd{(tmpdir,}\hlstr{'/chip'}\hlstd{),}\hlkwc{pattern}\hlstd{=}\hlstr{'*.bam$'}\hlstd{,}
                       \hlkwc{full.names}\hlstd{=T)}
\hlstd{bam.files}

\hlcom{# Design matrix}
\hlstd{design} \hlkwb{<-} \hlkwd{model.matrix}\hlstd{(}\hlopt{~}\hlkwd{factor}\hlstd{(cell.type))}
\hlkwd{colnames}\hlstd{(design)} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"intercept"}\hlstd{,} \hlstr{"cell.type"}\hlstd{)}
\hlstd{design}

\hlcom{# Parameters (PCR duplicates already removed and quality score filtered)}
\hlstd{param} \hlkwb{<-} \hlkwd{readParam}\hlstd{(}\hlkwc{dedup} \hlstd{= F)}
\hlstd{param}

\hlcom{# Estimating the average fragment length (rescaling all to 200bp)}
\hlstd{x} \hlkwb{=} \hlkwd{lapply}\hlstd{(bam.files,correlateReads,}\hlkwc{param}\hlstd{=param,}\hlkwc{max.dist}\hlstd{=}\hlnum{250}\hlstd{)}
\hlstd{multi.frag.lens} \hlkwb{=} \hlkwd{list}\hlstd{(}\hlkwd{unlist}\hlstd{(}\hlkwd{lapply}\hlstd{(x,maximizeCcf)),}\hlnum{200}\hlstd{)}
\hlstd{multi.frag.lens}

\hlcom{# Counting reads (for a window size of 250bp, for instance)}
\hlstd{data} \hlkwb{<-} \hlkwd{windowCounts}\hlstd{(bam.files,}\hlkwc{width} \hlstd{=} \hlnum{250}\hlstd{,}\hlkwc{ext} \hlstd{= multi.frag.lens,}
                     \hlkwc{param} \hlstd{= param,}\hlkwc{filter} \hlstd{=} \hlnum{20}\hlstd{)}
\hlstd{data}

\hlcom{# Filtering data}
\hlstd{data.large} \hlkwb{<-} \hlkwd{windowCounts}\hlstd{(bam.files,}\hlkwc{width}\hlstd{=}\hlnum{2500}\hlstd{,}\hlkwc{bin}\hlstd{=T,}\hlkwc{param}\hlstd{=param)}

\hlstd{bin.ab} \hlkwb{<-} \hlkwd{scaledAverage}\hlstd{(data.large,} \hlkwc{scale}\hlstd{=}\hlkwd{median}\hlstd{(}\hlkwd{getWidths}\hlstd{(data.large))}\hlopt{/}
                          \hlkwd{median}\hlstd{(}\hlkwd{getWidths}\hlstd{(data)))}

\hlstd{threshold} \hlkwb{<-} \hlkwd{median}\hlstd{(bin.ab)} \hlopt{+} \hlkwd{log2}\hlstd{(}\hlnum{2}\hlstd{)}

\hlstd{keep.global} \hlkwb{<-} \hlkwd{aveLogCPM}\hlstd{(}\hlkwd{asDGEList}\hlstd{(data))} \hlopt{>}  \hlstd{threshold}

\hlkwd{sum}\hlstd{(keep.global)}

\hlcom{# Creating filtered data}
\hlstd{filtered.data} \hlkwb{<-} \hlstd{data[keep.global,]}

\hlcom{# Testing for DB (assuming composition bias is negligble,}
\hlcom{# i.e. cell lines should exhibit a balanced number of DB regions) }

\hlstd{y} \hlkwb{<-} \hlkwd{DGEList}\hlstd{(}\hlkwd{assay}\hlstd{(filtered.data),} \hlkwc{lib.size} \hlstd{= filtered.data}\hlopt{$}\hlstd{totals)}
\hlstd{y}\hlopt{$}\hlstd{samples}\hlopt{$}\hlstd{norm.factors} \hlkwb{<-} \hlnum{1}
\hlstd{y}\hlopt{$}\hlstd{offset} \hlkwb{<-} \hlkwa{NULL}
\hlstd{y} \hlkwb{<-} \hlkwd{estimateDisp}\hlstd{(y, design)}
\hlstd{fit} \hlkwb{<-} \hlkwd{glmQLFit}\hlstd{(y, design,} \hlkwc{robust} \hlstd{=} \hlnum{TRUE}\hlstd{)}
\hlstd{out} \hlkwb{<-} \hlkwd{glmQLFTest}\hlstd{(fit,}\hlkwc{contrast} \hlstd{= contrast)}
\hlstd{tabres} \hlkwb{<-} \hlkwd{topTags}\hlstd{(out,} \hlkwd{nrow}\hlstd{(out))}\hlopt{$}\hlstd{table}
\hlstd{tabres} \hlkwb{<-} \hlstd{tabres[}\hlkwd{order}\hlstd{(}\hlkwd{as.integer}\hlstd{(}\hlkwd{rownames}\hlstd{(tabres))),]}

\hlstd{merged} \hlkwb{<-} \hlkwd{mergeWindows}\hlstd{(}\hlkwd{rowRanges}\hlstd{(filtered.data),} \hlkwc{tol}\hlstd{=tol,}
                       \hlkwc{max.width}\hlstd{=max.width)}
\hlstd{tabneg} \hlkwb{<-} \hlkwd{combineTests}\hlstd{(merged}\hlopt{$}\hlstd{id, tabres)}
\end{alltt}
\end{kframe}
\end{knitrout}
